---
title: "clustering method: k-means"
output: html_document
---
Purpose: To segment mall customers by their annual income and spending score using k-means clustering and identify target group for possible intervention(s) to increase their spending score.
The optimal number of clusters (k) is explored with the Elbow, average Silhouette and Gap statistic methods. Data source: Kaggle dataset (<https://www.kaggle.com/vjchoudhary7/customer-segmentation-tutorial-in-python/activity>).

#### Load required packages and data
```{r, message = FALSE}
library(ggplot2)
library(dplyr)
library(cluster)

cust <- read.csv("data.csv")
colnames(cust) <- c("cust_id", "gender", "age", "annual_income_000", "spend_score")
````

###### Elbow method
````{r}
kk <- 1:10
wss <- c()
for (i in kk){
  cluster_temp <- kmeans(cust[,c(4,5)], centers = i, nstart = 25)
  wss <- c(wss, cluster_temp$tot.withinss)
}
plot(wss, type = "b")
````
From the plot above, the optimal number of clusters appears to be at k=5.

###### Average Silhouette method
````{r}
kk <- 2:10
avg_silhou <- c()
for (i in kk){
  cluster_temp <- kmeans(cust[,c(4,5)], centers = i, nstart = 25)
  ss <- silhouette(cluster_temp$cluster, dist(cust[,c(4,5)]))
  avg_silhou <- c(avg_silhou, mean(ss[,3]))
}
plot(avg_silhou, type="b")
````
From plot above, maximum value of average silhouette occurs at no. of clusters, k = 4.

###### Gap statistic method
````{r}
set.seed(123)
gap_stat <- clusGap(cust[,c(4,5)], FUN = kmeans, nstart = 25, K.max = 10, B = 50)

# print result
print(gap_stat, method = "firstmax")
# plot result
fviz_gap_stat(gap_stat)
````
From plot above, maximum gap statistic occurs at no. of clusters, k = 5.

We will use `r kmeans` with `centers = 5` to segment the customers.
````{r}
set.seed(10)
cluster_5 <- kmeans(cust[,c(4,5)], centers = 5, nstart = 25)
# include the cluster number for each customer in dataset
cust <- cust %>% mutate(cluster5 = cluster_5$cluster)
# scatterplot
ggplot(cust, aes(x = annual_income_000, y = spend_score, colour = factor(cluster5))) +
  geom_point() +
  scale_color_discrete(name = "Clusters") +
  xlab("Annual income of customers (thousands)") +
  ylab("Spending score")
````
Interpretation of results:
Cluster 1: low annual income, low spending score
Cluster 2: average annual income, average score 
Cluster 3: high annual income, low score (interventions can be targeted at this group to increase their spending score)
Cluster 4: low annual income, high score
Cluster 5: high annual income, high score
